Stack Safety as a Security Property

### 2020-07-08 main.tex
Some notation for Section 6

\definition
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}

@@

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
  satisfying the stack integrity test $\mathit{int}_T ~ C
  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
  \mach ~ \mach'$ with regard to the original machine state. We write
  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.

@@

  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\mathit{stack}_T ~ d ~ [(\mach, \pol)] ~ \vs$, for any $d$ up to the size of
    $\vs$. (That is, a violation was detected $d$ levels into the stack.
    Elements are created and pushed to the variant stack on calls and popped
    from it on returns, so its size always reflects the depth of the program
    stack.)

  \item If the top variant of the stack (and only the top variant) returns,
    i.e., $\mathit{ret}(\head(\vs), \mach)$ holds, the property is safisfied for
    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
    stack: $\mathit{stack}_T ~ |\vs| ~ [(\mach, \pol)] ~ \vs$. \rb{Explain the
    use of this vs the return case below.}

  \item If
%
    (1) the machine takes a \emph{non-call, non-return step},
%
    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
    satisfied, and
%
    (3) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') : \MPT) ~
    \vs'$ is coinductively satisfied % at depth $d$
    %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
    %% \pol')$,
%
    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
    \pol') : \MPT) ~ vs$.
    %% \rb{omitting stack size condition}

  \item
%
    If (1) the machine takes a \emph{call step} and $\contour$ is the contour at
    the call,
%
    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
    satisfied,
%
    (3) the stack integrity test $\mathit{int}_T ~ \contour ~ \mach ~ \mach'$ is
    satisfied and the stack confidentiality test $\mathit{conf}_T ~ \mach ~
    \vm{\contour}$ is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') :
    \MPT) ~ (\vm{\contour} : \vs')$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
    \pol') : \MPT) ~ vs$.

@@

  \item
%
    If (1) the machine takes a \emph{return step} at the top of the variant
    stack (and only the top variant),
%
    (2) the variant stack test $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$ is
    satisfied,
%
    (3) the stack safety test $\mathit{stack}_T ~ d ~ ((\mach', \pol') : \MPT) ~
    \mathit{tail}(\vs)$ is coinductively satisfied
%
    {\em then} we conclude $\mathit{stack}_T ~ d ~ ((\mach, \pol) : (\mach',
    \pol') : \MPT) ~ vs$.

@@

\definition{Top-level stack testing property}
%
  \ldots wrap the coinductive stack testing property

>>>>

\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}
\newcommand*{\testInt}[3]{\ensuremath{\mathit{int}_T ~ #1 ~ #2 ~ #3}}
\newcommand*{\testConf}[2]{\ensuremath{\mathit{conf}_T ~ #1 ~ #2}}
\newcommand*{\testVS}[2]{\ensuremath{\mathit{\vs}_T ~ #1 ~ #2}}
\newcommand*{\testProp}[3]{\ensuremath{\mathit{stack}_T ~ #1 ~ #2 ~ #3}}

\definition
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \testInt{\contour}{\mach}{\mach'} \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}

@@

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
  and stack confidentiality test $\testConf{\mach}{\mach'}$
  with regard to the original machine state. We write
  this $\testVS{(\mach, \pol)}{\vs}$.

@@

  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\testProp{d}{[(\mach, \pol)]}{\vs}$, for any $d$ up to the size of
    $\vs$. (That is, a violation was detected $d$ levels into the stack.
    Elements are created and pushed to the variant stack on calls and popped
    from it on returns, so its size always reflects the depth of the program
    stack.)

  \item If the top variant of the stack (and only the top variant) returns,
    %i.e., $\mathit{ret}(\head(\vs), \mach)$ holds,
    the property is safisfied for
    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
    stack: $\testProp{|\vs|}{[(\mach, \pol)]}{\vs}$. \rb{Explain the
    use of this vs the return case below.}

  \item If
%
    (1) the machine takes a \emph{non-call, non-return step},
%
    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
    satisfied, and
%
    (3) the stack safety test $\testProp{d}{((\mach', \pol') : \MPT)}{\vs'}$
    is coinductively satisfied % at depth $d$
    %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
    %% \pol')$,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach', \pol') :
    \MPT)}{vs}$.
    %% \rb{omitting stack size condition}

  \item
%
    If (1) the machine takes a \emph{call step} and $\contour$ is the contour at
    the call,
%
    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
    satisfied,
%
    (3) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

@@

  \item
%
    If (1) the machine takes a \emph{return step}, so the top of the variant
    stack (and only the top variant) returns,
%
    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
    satisfied,
%
    (3) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{\mathit{tail}(\vs)}$ is coinductively satisfied
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

@@

\definition
%
  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
  satisfies the eager testing property at depth 1 with regard to the initial
  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
%
  \rb{I don't think the initial variant has been defined, but it's
    straightforward.}
